The importance of reproducibility in geographical research: the case of population density and the spread of COVID-19

### 2021-06-09 R0-Density-Reanalysis/R0-Density-Reanalysis.Rmd
Experiments with spatial filter in the outcome model of the Heckman selection model

```{r load-packages, include=FALSE, cache=FALSE}
library(censReg) # Package to estimate tobit model
library(kableExtra) # Package for creating tables
library(Matrix) # Used to convert list of spatial weights to sparse matrix; needed for spatialprobit
library(nlme) # Used to estimate 
library(patchwork)
library(r0density) # Data package
library(sampleSelection) # Package to estimate sampling selection models
library(sf) # Package to work with spatial information in simple features format
library(spatialprobit) # Used to estimate spatially autoregressive tobit
library(spdep) # Utilities to create spatial weights and work with spatial data
library(tidycensus) # Package to retrieve census information 
library(tidyverse) # Data carpentry and analysis tools
library(units) # Package to work with units of measurement
library(viridis) # Color palettes
```

@@

@@

@@

How relevant is the difference between these different model specifications? Figure \ref{fig:comparison-results} shows the relationship between density and $R_0$ implied by SWN Model 1 and Selection Model 2. The left panel of the figure shows the non-linear but monotonic relationship implied by SWN Model 1. The conclusion is that at higher densities, $R_0$ is _always_ higher. The right panel, in contrast, shows that, according to Selection Model 2, $R_0$ is zero when density is zero (as expected), and then it tends to increase at higher densities. This continues until a density of approximately 2.9 (1,000 people per sq.km). At higher densities than that $R_0$ begins to decline, and the relationship becomes negative at densities higher than approximately 5.7 (1,000 people per sq.km).

example_df <- data.frame(density = seq(0, 6.3, 0.01)) %>% 
  mutate(density_log = log(density),
         density_2 = density^2, 
         b_swn = model3$coefficients$fixed["density_log"] * density_log,
         b_selection = selection2$estimate[7] * density + selection2$estimate[8] * density_2)

density_swn + density_selection
```

@@

>>>>

```{r load-packages, include=FALSE, cache=FALSE}
library(adespatial) # To calculate Moran eigenvector maps
library(censReg) # Package to estimate tobit model
library(kableExtra) # Package for creating tables
library(Matrix) # Used to convert list of spatial weights to sparse matrix; needed for spatialprobit
library(nlme) # Used to estimate 
library(patchwork)
library(r0density) # Data package
library(sampleSelection) # Package to estimate sampling selection models
library(sf) # Package to work with spatial information in simple features format
library(spatialprobit) # Used to estimate spatially autoregressive tobit
library(spdep) # Utilities to create spatial weights and work with spatial data
library(tidycensus) # Package to retrieve census information 
library(tidyverse) # Data carpentry and analysis tools
library(units) # Package to work with units of measurement
library(viridis) # Color palettes
```

@@

county_listw <- nb2listw(county_nb, 
                        style="W", 
                        zero.policy = TRUE)

@@

```{r moran-eigenvector-maps, include=FALSE}
county_mem <- mem(county_listw)
```

@@

# Proceed with caution: Spatial effects ahead!

Words go here.

```{r selection-model-with-spatial-filter, include=FALSE}

# Refit model selection2 to begin 
selection3 <- selection(Ys ~ density + density_2 + private + hincome, R ~ density + density_2 + private + hincome, data = county_geo_clean)

# Extract the residuals of the model. These residuals have NAs where the outcome was not observed
res_selection3 <- residuals(selection3)

#Initialize a spatial filter:
SF <- numeric(length = length(res_selection3))

#Initialize tolerance, counter, and index:
tol <- 0;

while(tol < 0.25){
  #Initialize counter and index:
  count <- 0;
  SF_INDEX <- 0

  #Obtain filter to simulate the residuals
  for(i in 1:length(county_mem)){
    count <- count + 1
    junkSF <- data.frame(SF = SF, V = county_mem[, count])
    #remove columns with all zeros
    junkSF <- Filter(function(x)!all(x == 0), junkSF)
    #estimate model
    junkmod <- lm(res_selection3 ~ .,
                  data = cbind(res_selection3, junkSF))
    junkmod_summary <- summary(junkmod) # Summary of model; needed to extract the pval of the filter
    pvals<- junkmod_summary$coefficients[nrow(junkmod_summary$coefficients), 4]
    #pvals <- get_pvalue(junkmod)
    junkb <- coef(junkmod)
    if(pvals <= 0.10){
      SF_INDEX[count] <- count
      SF <- as.matrix(junkSF) %*% junkb[(0+2):length(junkb)]
      junkmod <- lm(res_selection3 ~ ., 
                    data = data.frame(res_selection3, SF))
    }
  }
  #SF_INDEX <- SF_INDEX[!is.na(SF_INDEX)]
  Filter1 <- SF # Spatial filter for residuals
  res_filter_cor <- cor(res_selection3, Filter1, use = "complete.obs")
  Filter_mt <- moran.test(Filter1, county_listw)
  tol <- Filter_mt$p.value
  if(tol < 0.25){
    selection3 <- selection(Ys ~ density + density_2 + private + hincome, 
                            R ~ density + density_2 + private + hincome + Filter1, 
                            data = county_geo_clean)
  }  
}
```

```{r prepare-selection3-results, include=FALSE}
selection3_me <- selection3 %>%
  summary()

selection3_me <- selection3_me$estimate %>%
  as.data.frame()

colnames(selection3_me) <- c("b", "se", "t", "p")

selection3_me <- selection3_me %>%  
  rownames_to_column(var = "Variable") %>% 
  transmute(Variable,
            b,
            ll = b - 1.96 * se, 
            ul = b + 1.96 * se)

colnames(selection3_me) <- c("Variable",
                                 "b1",
                                 "l1",
                                 "u1")

selection3_me <- selection3_me %>%
  transmute(Variable,
            b1 = ifelse(is.na(b1),
                        " ", 
                        round(b1, 3)),
            ci1 = ifelse(is.na(l1), 
                         " ",
                         paste0("[", round(l1, 3), ", ", round(u1, 3), "]")))
#%>%
#  arrange(Order) %>%
#  select(-Order)
```

```{r tabulate-sample-selection3-results, echo=FALSE}
selection3_me  %>%
  mutate(Variable = c("Intercept",
                      "Density (1,000 per sq.km)",
                      "Density squared",
                      "Percent of private transportation", 
                      "Median household income (10,000)",
                      "Intercept",
                      "Density (1,000 per sq.km)",
                      "Density squared",
                      "Percent of private transportation", 
                      "Median household income (\\$10,000)",
                      "Spatial filter",
                      "$\\sigma$",
                      "$\\rho$")) %>%
  kable("latex",
        digits = 3,
        booktabs = TRUE,
        escape = FALSE,
        col.names = c("Variable",
                      "$\\beta$", 
                      "95\\% CI"),
        align = c("lcccc"),
        caption = "\\label{tab:selection-results}Estimation results of sample selection model with spatial filter") %>%
  kable_styling(latex_options = c("scale_down")) %>% 
  pack_rows(group_label = "Sample Selection Model", 1, 5) %>%
  pack_rows(group_label = "Outcome Model", 6, 11) %>%
  add_header_above(c(" ", "Selection Model 3" = 2))
```

@@

# Discussion

How relevant are the difference between the different model specifications presented above? Figure \ref{fig:comparison-results} shows the relationship between density and $R_0$ implied by SWN Model 3, Selection Model 2, and Selection Model 3. The left panel of the figure shows the non-linear but monotonic relationship implied by SWN Model 1. The conclusion is that at higher densities, $R_0$ is _always_ higher. The right panel, in contrast, shows that, according to Selection Model 2, $R_0$ is zero when density is zero (as expected), and then it tends to increase at higher densities. This continues until a density of approximately 2.9 (1,000 people per sq.km). At higher densities than that $R_0$ begins to decline, and the relationship becomes negative at densities higher than approximately 5.7 (1,000 people per sq.km).

example_df <- data.frame(density = seq(0, 6.3, 0.01)) %>% 
  mutate(density_log = log(density),
         density_2 = density^2, 
         b_swn = model3$coefficients$fixed["density_log"] * density_log,
         b_selection = selection2$estimate[7] * density + selection2$estimate[8] * density_2,
         b_selection3 = selection3$estimate[7] * density + selection3$estimate[8] * density_2)

# Plot of coefficient of density to R_0  
density_selection3 <- ggplot(data = example_df) + 
  geom_line(aes(x = density, 
                y = b_selection3)) +
  ggtitle("Selection Model 3") +
  ylab(expression(paste("E[", R[0], "]"))) +
  theme_minimal()

density_swn + density_selection + density_selection3
```

### 2021-06-09 R0-Density-Reanalysis/R0-Density-Reanalysis.tex
Experiments with spatial filter in the outcome model of the Heckman selection model

How relevant is the difference between these different model
specifications? Figure \ref{fig:comparison-results} shows the
relationship between density and \(R_0\) implied by SWN Model 1 and
Selection Model 2. The left panel of the figure shows the non-linear but
monotonic relationship implied by SWN Model 1. The conclusion is that at
higher densities, \(R_0\) is \emph{always} higher. The right panel, in
contrast, shows that, according to Selection Model 2, \(R_0\) is zero
when density is zero (as expected), and then it tends to increase at
higher densities. This continues until a density of approximately 2.9
(1,000 people per sq.km). At higher densities than that \(R_0\) begins
to decline, and the relationship becomes negative at densities higher
than approximately 5.7 (1,000 people per sq.km).

>>>>

\hypertarget{proceed-with-caution-spatial-effects-ahead}{%
\section{Proceed with caution: Spatial effects
ahead!}\label{proceed-with-caution-spatial-effects-ahead}}

Words go here.

\caption{\label{tab:tabulate-sample-selection3-results}\label{tab:selection-results}Estimation results of sample selection model with spatial filter}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Selection Model 3} \\
\cmidrule(l{3pt}r{3pt}){2-3}
Variable & $\beta$ & 95\% CI\\
\midrule
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Sample Selection Model}}\\
\hspace{1em}Intercept & -7.304 & [-8.346, -6.262]\\
\hspace{1em}Density (1,000 per sq.km) & 2.445 & [2.089, 2.802]\\
\hspace{1em}Density squared & -0.380 & [-0.468, -0.292]\\
\hspace{1em}Percent of private transportation & 0.056 & [0.046, 0.067]\\
\hspace{1em}Median household income (10,000) & 0.318 & [0.28, 0.356]\\
\addlinespace[0.3em]
\multicolumn{3}{l}{\textbf{Outcome Model}}\\
\hspace{1em}Intercept & 2.563 & [2.497, 2.629]\\
\hspace{1em}Density (1,000 per sq.km) & 0.760 & [0.746, 0.774]\\
\hspace{1em}Density squared & -0.133 & [-0.135, -0.13]\\
\hspace{1em}Percent of private transportation & -0.011 & [-0.012, -0.011]\\
\hspace{1em}Median household income (\$10,000) & 0.002 & [-0.001, 0.004]\\
\hspace{1em}Spatial filter & 1.000 & [0.998, 1.001]\\
$\sigma$ & 0.017 & [0.015, 0.019]\\
$\rho$ & -0.304 & [-0.957, 0.349]\\
\bottomrule
\end{tabular}}
\end{table}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

How relevant are the difference between the different model
specifications presented above? Figure \ref{fig:comparison-results}
shows the relationship between density and \(R_0\) implied by SWN Model
3, Selection Model 2, and Selection Model 3. The left panel of the
figure shows the non-linear but monotonic relationship implied by SWN
Model 1. The conclusion is that at higher densities, \(R_0\) is
\emph{always} higher. The right panel, in contrast, shows that,
according to Selection Model 2, \(R_0\) is zero when density is zero (as
expected), and then it tends to increase at higher densities. This
continues until a density of approximately 2.9 (1,000 people per sq.km).
At higher densities than that \(R_0\) begins to decline, and the
relationship becomes negative at densities higher than approximately 5.7
(1,000 people per sq.km).
